An Approach for the Reduction of Boundary Effects in Time-Frequency Representations

### 2020-06-01 Notes/BoundEffRed.tex
paper
updates

\begin{figure}
\centering
%\includegraphics[width=.48\textwidth]{PPGsig.eps}
\includegraphics[width=.48\textwidth]{SSTintro.eps}
\includegraphics[width=.48\textwidth]{zoomSSTintro.eps}
\caption{Synchrosqueezing transform of a PPG (top) with a zoom on its right boundary (bottom). }
\label{fig:ex.intro}
\end{figure}

@@

Various improvements can be considered to make the algorithm more robust to certain situations. In particular, we have noticed that when the observed signal does not oscillate, the forecasting is no longer relevant, and outlier time-frequency representations are obtained. A preliminary step should then be added to the algorithm to select the portions of the signal that exhibit oscillations, in order to reject the portions for which accurate time-frequency analysis is practically useless. A more fundamental improvement would be to perform the prediction step directly in the time-frequency plane, with the aim of specifically predicting the evolution of each instantaneous frequency. This would lead to seeing a time-frequency representation as a set of interdependent time series (corresponding to each frequency channel).

>>>>

\begin{figure}
\centering
\includegraphics[width=.48\textwidth]{zoomPPGsig.eps}
\includegraphics[width=.48\textwidth]{SSTintro.eps}
\includegraphics[width=.48\textwidth]{zoomSSTintro.eps}
\caption{Zoom on a PPG signal (top), with its synchrosqueezing transform (middle) and a zoom on the right boundary of the SST (bottom). The window length is 12~sec. }
\label{fig:ex.intro}
\end{figure}

@@

Various improvements can be considered to make the algorithm more robust to certain situations. In particular, we have noticed that when the observed signal does not oscillate, the forecasting is no longer relevant, and outlier time-frequency representations are obtained. A preliminary step should then be added to the algorithm to select the portions of the signal that exhibit oscillations, in order to reject the portions for which accurate time-frequency analysis is practically useless. A more fundamental improvement would be to perform the prediction step directly in the time-frequency plane, with the aim of specifically predicting the evolution of each instantaneous frequency. This would lead to seeing a time-frequency representation as a set of interdependent time series (corresponding to each frequency band).

### 2020-06-01 Notes/Supplementary material/SupMat.tex
updates

\usepackage{graphicx}
\usepackage{bbm}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{type1cm,eso-pic,color}
\usepackage{mathrsfs} 
\usepackage{mathpazo}
\usepackage[scaled=.95]{helvet}
\usepackage{courier}
\usepackage{xr}
\usepackage{xspace}
\usepackage{etoolbox}
\usepackage[shortlabels]{enumitem}
\usepackage{multirow}
\usepackage{dsfont}
\usepackage{hhline}
\usepackage{algorithm,algorithmic}
\usepackage{xr}
\externaldocument[I-]{../BoundEffRed}

@@

\setcounter{equation}{8}
\setcounter{figure}{3}

@@

\subsection{Notations}
We have:
\begin{align}
\dfrac1K\bX\bX^T &= \underbrace{\dfrac1K\bZ\bZ^T+\sigma^2\bI}_{\defeq S^{(0)}} + \bE^{(0)} \\
\dfrac1K\bY\bX^T &= \underbrace{\dfrac1K\bZ'\bZ^T +\sigma^2\bD}_{\defeq S^{(1)}}+ \bE^{(1)} \ ,
\end{align}
where $\bE^{(a)} = \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$ with:
\[
\bE_1^{(a)}[m,m'] = \dfrac1K\sum_{k=0}^{K-1} \bz[N_0+m+a+k]\bw[N_0+m'+k] + \bw[N_0+m+a+k]\bz[N_0+m'+k]\ ,
\]
and
\[
\bE_2^{(a)}[m,m'] =  \dfrac1K\sum_{k=0}^{K-1} \bw[N_0+m+a+k]\bw[N_0+m'+k] - \delta_{(m+a)m'}\ ,
\]
with $a\in\{0,1\}$.

@@

\subsection{Study of $\bh^{(\ell)}$}
The randomness of $\bh^{(\ell)}$ is completely originating from the error matrices. Besides, notice that the first $M-1$ rows in $\bE^{(1)}$ are the same as the last $M-1$ rows of $\bE^{(0)}$. Thus, we gather all the sources of randomness into an vector $\bg\in\RR^{M(M+1)}$ defined as:
\[
\bg = \vec\left(
\begin{pmatrix}
\bE^{(0)} \\
\be_M^T\bE^{(1)}
\end{pmatrix}
\right)\ .
\]
Here, "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another. Then, by definition, we have:
\[
\bg = \sigma\bg_1 + \sigma^2\bg_2\ ,
\]
where:
\begin{align*}
\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right) \\
\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \ .
\end{align*}
with $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[0] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[0] \end{pmatrix}$. Then, $\bg_1$ is a Gaussian random vector as a linear combination of Gaussian random vectors. Moreover, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$. Combining these two results gives the following result:
\[
\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ ,
\]
where $\bGamma_0=\EE\left\{\bg\bg^T\right\}$ is a covariance matrix.

>>>>

\usepackage{graphicx}
\usepackage{bbm}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{type1cm,eso-pic,color}
\usepackage{mathrsfs} 
\usepackage{mathpazo}
\usepackage[scaled=.95]{helvet}
\usepackage{courier}
\usepackage{xr}
\usepackage{xspace}
\usepackage{etoolbox}
\usepackage[shortlabels]{enumitem}
\usepackage{multirow}
\usepackage{dsfont}
%\usepackage{hhline}
%\usepackage{algorithm,algorithmic}

\usepackage{xr}
\externaldocument[I-]{../BoundEffRed}

@@

@@

\subsection{Notations}
By definition the matrix $\bX$ and $\bY$ and because of the noisy signal model~\eqref{eq:model.noise}, we have:
\begin{align}
\dfrac1K\bX\bX^T &= \underbrace{\dfrac1K\bZ\bZ^T+\sigma^2\bI}_{\defeq S^{(0)}} + \bE^{(0)} \\
\dfrac1K\bY\bX^T &= \underbrace{\dfrac1K\bZ'\bZ^T +\sigma^2\bD}_{\defeq S^{(1)}}+ \bE^{(1)} \ ,
\end{align}
where $\bE^{(a)} = \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$, with:
\[
\bE_1^{(a)}[m,m'] = \dfrac1K\sum_{k=0}^{K-1} \bz[N_0+m+a+k]\bw[N_0+m'+k] + \bw[N_0+m+a+k]\bz[N_0+m'+k]\ ,
\]
and
\[
\bE_2^{(a)}[m,m'] =  \dfrac1K\sum_{k=0}^{K-1} \bw[N_0+m+a+k]\bw[N_0+m'+k] - \delta_{(m+a)m'}\ ,
\]
with $a\in\{0,1\}$.

@@

\subsection{Study of $\bh^{(\ell)}$}
The randomness of $\bh^{(\ell)}$ is completely originating from the error matrices. Besides, notice that the first $M-1$ rows in $\bE^{(1)}$ are equals to the last $M-1$ rows of $\bE^{(0)}$. Thus, we gather all the sources of randomness into an vector $\bg\in\RR^{M(M+1)}$, containing $M$ rows and defined as:
\[
\bg = \vec\left(
\begin{pmatrix}
\bE^{(0)} \\
\be_M^T\bE^{(1)}
\end{pmatrix}
\right)\ .
\]
Here, "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another. Then, by definition, we have:
\[
\bg = \sigma\bg_1 + \sigma^2\bg_2\ ,
\]
where:
\begin{align*}
\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right) \\
\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \ .
\end{align*}
with $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[0] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[0] \end{pmatrix}$. Then, $\bg_1$ is a Gaussian random vector because it is a linear combination of Gaussian random vectors. Moreover, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$. Combining these two results gives the following result:
\[
\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ ,
\]
where $\bGamma_0=\EE\left\{\bg\bg^T\right\}$ is a covariance matrix.

### 2020-06-01 Notes/results.tex
updates

To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of the Algorithm~\ref{alg:boundary} with reference forecasting algorithm that could be used for extending such signals. These methods are:
\begin{itemize}
\item The \emph{Extended Dynamic Mode Decomposition (EDMD)} has been developped by Williams \etal~\cite{Williams15data}. The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows to catch dynamic of nonlinear systems~\cite{Korda18linear}.
\item The \emph{Gaussian Process Regression (GPR)}~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model. That one is based on the Gaussian process structure, and therefore offer more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
\item The \emph{Trigonometric , Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS)} method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, a seasonal and an ARMA components, with a specific dynamic for the seasonal component. This model demands the estimation of numerous parameters and, by implication, may be slow. 
\end{itemize}

@@

\begin{table}
\centering
\caption{AHM signal. Performance of the extension methods.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
   \cline{2-3} & Mean & Standard deviation & \\
   \hhline{|=#=|=|=|}
   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.15$ \\
   \hline
   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.53$\\
   \hline
   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.33$ \\
   \hline
   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.12$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}

@@

\begin{table}
\centering
\caption{Respiratory signal. Averaged OTD of the boundary effects reduction methods on diverse representations.}
\begin{tabular}{|>{\centering}m{35pt}||c|c|c|c|}
  \hline
   \multirow{2}{35pt}{\centering Extension method} & \multicolumn{4}{c|}{Time-Frequency Representation} \\
   \cline{2-5}
      & STFT & SST & RS & ConceFT \\
   \hhline{|=#=|=|=|=|}
   Without extension & $2.16\!\times\! 10^{-2}$ & $5.26\!\times\! 10^{-3}$ & $3.07\!\times\! 10^{-2}$ & $1.41\!\times\! 10^{-2}$ \\
   \hline
   {\sf SigExt} & $1.72\!\times\! 10^{-2}$ & $4.00\!\times\! 10^{-3}$ & $2.43\!\times\! 10^{-2}$ & $1.12\!\times\! 10^{-2}$ \\
   \hline
   EDMD & $1.75\!\times\! 10^{-2}$ & $4.23\!\times\! 10^{-3}$ & $2.45\!\times\! 10^{-2}$ & $1.14\!\times\! 10^{-2}$ \\
   \hline
   GPR & $1.84\!\times\! 10^{-2}$ & $4.10\!\times\! 10^{-3}$ & $2.48\!\times\! 10^{-2}$ & $1.10\!\times\! 10^{-2}$ \\
   \hline
\end{tabular}
\label{tab:otd.tho}
\end{table}

>>>>

To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AM-FM, we compare the performance of the Algorithm~\ref{alg:boundary} with reference forecasting algorithm that could be used for extending such signals. These methods are:
\begin{itemize}
\item The \emph{Extended Dynamic Mode Decomposition (EDMD)} has been developped by Williams \etal~\cite{Williams15data}. The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows to catch dynamic of nonlinear systems~\cite{Korda18linear}.
\item The \emph{Gaussian Process Regression (GPR)}~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model. That one is based on the Gaussian process structure, and therefore offer more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
\item The \emph{Trigonometric , Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS)} method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, a seasonal and an ARMA components, with a specific dynamic for the seasonal component. This model demands the estimation of numerous parameters and, by implication, may be slow. 
\end{itemize}

@@

\begin{table}
\centering
\caption{AM-FM signal. Performance of the extension methods.}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
   \cline{2-3} & Mean & Standard deviation & \\
   \hhline{|=#=|=|=|}
   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.15$ \\
   \hline
   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.53$\\
   \hline
   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.33$ \\
   \hline
   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.12$ \\
   \hline
\end{tabular}
\label{tab:mse.sine}
\end{table}

@@

\begin{table}
\centering
\caption{Respiratory signal. Averaged OTD of the boundary effects reduction methods on diverse representations.}
\begin{tabular}{|>{\centering}m{35pt}||c|c|c|c|}
  \hline
   \multirow{2}{35pt}{\centering Extension method} & \multicolumn{4}{c|}{Time-Frequency Representation} \\
   \cline{2-5}
      & STFT & SST & RS & ConceFT \\
   \hhline{|=#=|=|=|=|}
   Without extension & $2.16\!\times\! 10^{-2}$ & $5.26\!\times\! 10^{-3}$ & $3.07\!\times\! 10^{-2}$ & $1.41\!\times\! 10^{-2}$ \\
   \hline
   {\sf SigExt} & $1.72\!\times\! 10^{-2}$ & $4.00\!\times\! 10^{-3}$ & $2.43\!\times\! 10^{-2}$ & $1.12\!\times\! 10^{-2}$ \\
   \hline
   EDMD & $1.75\!\times\! 10^{-2}$ & $4.23\!\times\! 10^{-3}$ & $2.45\!\times\! 10^{-2}$ & $1.09\!\times\! 10^{-2}$ \\
   \hline
   GPR & $1.84\!\times\! 10^{-2}$ & $4.10\!\times\! 10^{-3}$ & $2.48\!\times\! 10^{-2}$ & $1.24\!\times\! 10^{-2}$ \\
   \hline
\end{tabular}
\label{tab:otd.tho}
\end{table}
